Results of two multi-chord stellar occultations by dwarf planet (1) Ceres

### 2017-08-16 Ceres/Old_versions/ArtivoV9/CeresV9.tex
mod 2017 ago16

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage[T1]{fontenc}
\usepackage{prettyref}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{epstopdf}
\usepackage{color}
\usepackage[breaklinks=true]{hyperref}

@@

\title[Results of two stellar occultations by Ceres]{Results of two multi-chord stellar occultations by dwarf planet (1) Ceres}
\author[A. R. Gomes-J\'unior, B. L. Giacchini, F. Braga-Ribas et al.]{A. R. Gomes-J\'unior$^{1,}$\thanks{E-mail: altair08@astro.ufrj.br},
B. L. Giacchini$^{2,3,4}$,%\thanks{E-mail: breno@cbpf.br},
F. Braga-Ribas$^{5, 6, \dag}$,
M. Assafin$^{1, \dag, \ddag}$,\newauthor
R. Vieira-Martins$^{1,5, \dag, \ddag}$,
J.I.B. Camargo$^{5, \dag}$,
B. Sicardy$^{7}$,
B. Timerson$^{4}$, 
T. George$^{4}$, \newauthor
J. Broughton$^{8}$, 
T. Blank$^{4}$,
G. Benedetti-Rossi$^{5}$, 
J. Brooks$^{4}$,  
R. F. Dantowitz$^{9}$,  \newauthor
D. W. Dunham$^{4}$, 
J. B. Dunham$^{4}$, 
C. K. Ellington$^{4}$,
M. Emilio$^{10}$,
F.R. Herpich$^{11}$,  \newauthor
C. Jacques$^{3, 12}$,
P. D. Maley$^{4,13}$,
L. Mehret$^{10}$,
A.J.T. Mello$^{14}$,
A.C. Milone$^{15}$,\newauthor
E. Pimentel$^{3, 12}$,  
W. Schoenell$^{11}$,
N. S. Weber$^{9}$
\\
$^{1}$Observat\'orio do Valongo/UFRJ, Ladeira Pedro Ant\^onio 43,
CEP 20.080-090 Rio de Janeiro - RJ, Brazil\\
$^{2}$Centro Brasileiro de Pesquisas F\'isicas, Rua Dr. Xavier Sigaud 150, Rio de Janeiro  22290-180, Brazil\\
$^{3}$Se\c{c}\~ao de Oculta\c{c}\~oes/REA-Brasil, Belo Horizonte, Brazil\\
$^{4}$International Occultation Timing Association, P.O. Box 131034, Houston, TX 77219-1034, USA\\
$^{5}$Observat\'orio Nacional/MCTI, R. General Jos\'e Cristino 77, CEP 20921-400 Rio de Janeiro - RJ, Brazil\\
$^{6}$Federal University of Technology - Paran\'a (UTFPR / DAFIS), Rua Sete de Setembro, 3165, CEP 80230-901, Curitiba, PR, Brazil\\
$^{7}$LESIA, Observatoire de Paris, CNRS UMR 8109, Universit\'{e} Pierre et Marie Curie, Universit\'{e} Paris-Diderot, 5 place Jules Janssen,\\ F-92195 Meudon Cedex, France\\
$^{8}$RASNZ Occultation Section, P.O. Box 3181, Wellington, New Zealand\\
$^{9}$Clay Center Observatory at Dexter Southfield, 20 Newton Street, Brookline, MA 02445, USA\\
$^{10}$Universidade Estadual de Ponta Grossa, Ponta Grossa, Brazil\\
$^{11}$Universidade Federal de Santa Catarina, Florian\'opolis, Brazil\\
$^{12}$Centro de Estudos Astron\^omicos de Minas Gerais, Belo Horizonte, Brazil\\
$^{13}$NASA Johnson Space Center Astronomical Society, Houston, Texas, USA\\
$^{14}$Federal University of Technology - Paran\'a (UTFPR / DAELT), Rua Sete de Setembro, 3165, CEP 80230-901, Curitiba, PR, Brazil\\
$^{15}$Instituto Nacional de Pesquisas Espaciais, S\~ao Jos\'e dos Campos, Brazil\\
$^\dag$ Associated to Laborat\'{o}rio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, CEP 20921-400,\\ Rio de Janeiro, Brazil\\
$^\ddag$Affiliated researcher at Observatoire de Paris/IMCCE, 77 Avenue Denfert Rochereau 75014 Paris, France
}
\begin{document}

@@

\begin{abstract}
We report the results of two multi-chord stellar occultations by the dwarf planet (1) Ceres that were observed  from Brazil on 2010 August 17, and from the USA on 2013 October 25. Four positive detections were obtained for the 2010 occultation, and nine for the 2013 occultation. Elliptical models were adjusted to the observed chords to obtain Ceres' size and shape. Two limb fitting solutions were studied for each event. The first one is a nominal solution with an indeterminate polar aspect angle. The second one was constrained by the pole coordinates as given by Drummond et al. Assuming a Maclaurin spheroid, we determine an equatorial diameter of 972 $\pm$ 6 km and an apparent oblateness of $0.08 \pm 0.03$ as our best solution. These results are compared to all available size and shape determinations for Ceres made so far, and shall be confirmed by the NASA's \textit{Dawn} space mission.
\end{abstract}

@@

This work is organized as follows. In Sections 2 and 3 we analyse the 2010 and the 2013 occultations, respectively. In Section 4 we give the geocentric positions of Ceres derived from the occultations. The comparison of our results to those in the literature is carried out in Section 5.

@@

As predicted, on 2010 August 17 Ceres occulted the star TYC 6833-163-1 (UCAC4 313-111823), which has magnitude $V = 11.55$ and has the ICRS position, based on UCAC4 catalogue \citep{Zacharias2013}, to the date of occultation:

%abaixo (UCAC2 20678706),
\begin{equation}
\left\{ 
  \begin{array}{l l}
    \alpha = 17^{\rmn{h}}18^{\rmn{m}}29\fs0085\\
    \delta = -27\degr 26\arcmin 38\farcs 867
  \end{array}
\right.
\label{Eq: 2010-coord}
\end{equation}

@@

Observations were carried out in Brazil from five different sites as displayed in Table \ref{Tab: obs-2010} and Fig. \ref{Fig: Ceres-2010-map}. The occultation was detected from four of them. The southern-most one (UFSC) had a negative chord. From the positive sites, the one named INPE started observing after the event was already in progress, due to technical difficulties, thus providing only the star's reappearance time; the other three recorded the whole phenomenon.

A remarkable circumstance of this event was the low velocity of Ceres: only 3.9 km s$^{-1}$ in the plane of the sky. Therefore, even exposures of a few seconds would translate in a relevant spatial resolution.

@@

\begin{table*}
 \centering
 \begin{minipage}{140mm}
  \caption{Circumstances of observation for all observing stations of the 2010 event. \label{Tab: obs-2010}}
  \begin{tabular}{@{}lccccc}
  \hline
          & Longitude & Telescope & Exposure & Result &   \\
     Site & Latitude  & Aperture  & Cycle time & Ingress & Observer\\          
          & Height    & f-ratio & Camera & Egress    & \\          
\hline
 Belo Horizonte & 43$\degr$59\arcmin 51\farcs1~W & LX200 & 5 s & Positive & C. Jacques  \\
 CEAMIG-REA &19$\degr$49\arcmin 49\farcs0~S & 31 cm & 12 s & 22:39:03.9 $\pm$ 0.6 s &  E. Pimentel \\
            & 825 m       &  f/10         &   SBIG ST10        & 22:40:20 $\pm$ 5 s &   \\
 & & & & & \\
 Pico dos Dias    & 45$\degr$34\arcmin45\farcs1~W &  Zeiss     & 1 s & Positive & J. I. B. Camargo \\
 LNA    &22$\degr$32\arcmin03\farcs7~S & 60 cm & 2 s & 22:37:30.3 $\pm$ 0.6 s &  G. B. Rossi \\
            & 1864 m     & f/12.5          &    Andor Ikon       & 22:41:55.3 $\pm$ 0.7 s &              \\
 & & & & & \\
 S\~ao Jos\'e dos       & 45$\degr$51\arcmin44\farcs0~W &  C11  & 2 s & Egress only & A. C. Milone\\
 Campos       &23$\degr$12\arcmin33\farcs0~S & 28 cm & 5 s & Start obs.: 22:39:44 & T. Maldonado\\
 INPE      & 620 m      & f/6.3          & SBIG ST7     & 22:42:03.0 $\pm$ 0.2 s & M. Okada    \\
 & & & & & \\
 Ponta Grossa       & 50$\degr$05\arcmin56\farcs0~W & RCX 400 &30 s & Positive & M. Emilio   \\
 UEPG       &25$\degr$05\arcmin22\farcs2~S & 40 cm & 52 s & 22:37:17 $\pm$ 13 s & L. Mehret   \\
            & 910 m      & f/8          & SBIG STL6E     & 22:39:56 $\pm$ 13 s & \\
 & & & & & \\
 Florian\'opolis       & 48$\degr$31\arcmin20\farcs5~W &   C11      & 3 s & No occultation  & W. Schoenell\\
 UFSC       &27$\degr$36\arcmin12\farcs3~S & 28 cm & 6 s&  Start obs.: 21:49:27 & A. J. T. Mello\\
            & 20 m       & f/10          &  SBIG ST7      &   End obs.: 22:51:21      & F. R. Herpich \\
\hline
\end{tabular}
\end{minipage}
\end{table*}

\subsection{Light curves \label{Sec: light-curve-2010}}

@@

The ingress (disappearance) and egress (reappearance) instants of the occultation were obtained for each light curve by fitting a square-well model convoluted with the Fresnel diffraction, the CCD bandwidth, the stellar apparent diameter, and the applied finite exposure time; see \cite{Widemann2009} and BR13. %\cite{BragaRibas2013}.
%The Fresnel scale ($F = \sqrt{\lambda D/2}$) for the present Ceres' geocentric distance $D = 2.29 \text{ au} = 3.42 \times 10^{8}$~km is 0.3~km for a typical wavelength of $\lambda = 0.65$~$\micron$. The star diameter is estimated using the formulae of \cite{vanBelle1999}. Its $B$, $V$, and $K$ apparent magnitudes are 12.00, 11.78 and 9.18, respectively, in the Tycho catalog \citep{Hog2000}. This yields a diameter of about 0.11~km projected at Ceres' distance. 
The smallest integration time used in the positive observations was 1.0~s, which translates to almost 3.9 km in the celestial plane. Therefore, the error on the time determination of the ingress and egress is largely dominated by the integration times, not by Fresnel diffraction or star diameter, which are both on the order of a few hundred meters for this event.
The occultation data fit consists in minimizing a classical $\chi^{2}$ function for each light curve, as described in \cite{Sicardy2011} and BR13.\ The free parameter to adjust is the ingress or egress instant, which provides the minimum value of $\chi^{2}$, denoted as $\chi^{2}_{min}$. The best fittings to the 2010 occultation light curves are shown in Fig. \ref{Fig: Ceres-2010-curves}, and the derived occultation times are listed in Table~\ref{Tab: obs-2010}.
%The occultation fits consist in minimizing a classical $\chi^{2}$ function for each light curve, as described in \cite{Sicardy2011} and BR13.\ The free parameter to adjust is the ingress (disappearance) or egress (reappearance) time $t_{occ}$, which provides the minimum value of $\chi^{2}$, denoted as $\chi^{2}_{min}$. The best fittings to the 2010 occultation light curves are shown in Fig. \ref{Fig: Ceres-2010-curves}, and the derived occultation times are listed in Table~\ref{Tab: obs-2010}.

@@

We adopt an elliptic model for the limb profile, resulting from the projection of an oblate spheroid onto the sky plane. This choice is supported by the work of \cite{Drummond2014}, by means of direct imaging of Ceres. Hence, we have $N=7$ chord extremities to adjust the $M=5$ parameters which define an ellipse: apparent semimajor and semiminor axis ($a^\prime$ and $b^\prime$, respectively), position angle $P$ of its semiminor axis and the position $(f_c,g_c)$ of its centre with respect to the occulted star. 
Of course, the apparent semimajor axis {$a^\prime$} is equivalent to the equatorial radius $R_{equa}$ of the ellipsoid.
The coordinates $f_{c}$ and $g_{c}$, in kilometres, are calculated using the JPL\#33 Ceres' ephemeris \citep{Giorgini1996} and the occulted star's position. They are positive toward the local celestial east and north, respectively. The position angle $P$ is counted positively from the direction of local celestial north to celestial east. The apparent oblateness can be defined by $\epsilon^\prime = 1 - (b^\prime/a^\prime)$. The best-fitting solution is obtained minimizing a reduced $\chi^{2}_{r}$ function, where we define the number of degrees of freedom of the problem as $\mathcal{N} \equiv N - M$.  All the procedures that allow the determination of the error bars of the physical parameters can be found on BR13.

@@

Two possible solutions were considered for the limb fitting. The first, which we call nominal solution, consists of determining the five parameters that characterize an ellipse from the seven observed contacts. As we shortly show, it led to a rather large uncertainty on the position angle. Furthermore, the nominal solution alone is not capable of returning the true oblateness, which can be evaluated through equation 2 of BR13  provided that the aspect angle $\zeta$ is known.% or constrained by some physical hypothesis.

@@

\begin{table*}
 \centering
 \begin{minipage}{140mm}
  \caption{Results of limb fitting to the data of the 2010 and 2013 events.\label{Tab: results}}
  \begin{tabular}{@{}lcccc}
  \hline
     Solution & 2010/Nominal & \textbf{2010/Pole-constrained} & 2013/Nominal & 2013/Pole-constrained \\
\hline
Equatorial diameter (km) & 982 $\pm$ 14 & \textbf{972 $\pm$ 6}  & 971 $\pm$ 7  & 971 $\pm$ 7\\
%Equatorial radius (km) & 491 $\pm$ 7  & \textbf{486 $\pm$ 3}  & 485.5 $\pm$ 3.5 & 485.5 $\pm$ 3.5\\
True oblateness        & 0.08 $\pm$ 0.03 & \textbf{0.08 $\pm$ 0.03} & 0.08 $\pm$ 0.04 & 0.08 $\pm$ 0.04\\
Position angle (deg)   & 5 $\pm$ 10    & \textbf{12 $\pm$ 3} (*)& 22 $\pm$ 5    & 25 $\pm$ 3 (*)\\
$f_c$ (km)             & 97 $\pm$ 9   & \textbf{102 $\pm$ 5}   & 77 $\pm$ 6    & 78 $\pm$ 6\\
$g_c$ (km)             & 16 $\pm$ 15  & \textbf{21 $\pm$ 11}  & 13 $\pm$ 16   & 13 $\pm$ 16\\
$\chi^2_{r,min}$       & 0.24          &  \textbf{0.42}         & 1.27          & 1.27\\
\hline
\end{tabular}
\textbf{Notes.} In bold we highlight our best solution. Error bars are at 1-sigma level. The polar diameter ($D_{pol}$) can be easily calculated from $D_{pol}=D_{equa}(1 - \epsilon)$.
(*) Position angle derived from Ceres' rotational pole coordinates determined by \cite{Drummond2014}.
\end{minipage}
\end{table*}

\begin{figure}
\includegraphics[scale=0.37, angle=-90]{figures/Ceres_2010_sphere.eps} 
\caption{The best elliptical fit for the occultation chords for the event of 2010 using the times from Table~\ref{Tab: obs-2010} and the pole-constrained solution. The arrow indicates the direction of motion, blue lines are the observed chords, the red segments are the ingress, egress and mid-occultation error bars at 1-$\sigma$ level. \label{Fig:Ceres-2010-body}}
\end{figure}

@@

The event which took place on 2013 October 25 involved the star TYC 865-911-1 (UCAC4 496-058191), of magnitude $V = 10.05$. Based on UCAC4 \citep{Zacharias2013}, the ICRS position to the date of occultation is:
%
\begin{equation}
\left\{ 
  \begin{array}{l l}
    \alpha = 11^{\rmn{h}}57^{\rmn{m}}52\fs7641\\
    \delta = +09\degr 07\arcmin 49\farcs835
  \end{array}
\right.
\end{equation}
%
The occultation could only be visible in the United States, before dawn, as depicted in Fig. \ref{Fig: Ceres-2013-map}.% The shadow would then cross the Atlantic and Africa, but during daylight time.

@@

Nine positive chords were obtained by the variety of instruments listed in Table \ref{Tab: obs-2013}. Each station was equipped with a video camera with negligible readout times. This was of particular importance, since in this event Ceres' shadow speed was 42.6 km s$^{-1}$, much faster than in the 2010 event.

@@

\subsection{Light curves}

@@

The immersion recorded at Owings also seems to be shifted (delayed) with respect to the nearby chords (see Fig. \ref{Fig: Ceres-2013-body}). This chord, actually, has roughly the same length of Mechanicsville's, despite the fact that they are separated by about 100 km. Differently from Varina, though, this station had time stamps inserted in each frame of the video, which makes it more unlikely to justify an eventual timing issue. Another possible explanation for the apparent problem of the times of this chord is the determination of the ingress and egress instants in the light curve analysis, which could have been affected by noise. Finally, the delay during the ingress could be caused by a relief feature in Ceres; we shall soon return to this hypothesis.

@@

In the hypothesis that the delay observed in the immersion at Owings could be associated to a limb topography feature, the recorded contact would then correspond to a negative elevation of 31 $\pm$ 4 km with respect to the best-fitting ellipse. Theoretical models, however, predict that reliefs in Ceres should not be higher than about 10--20 km \citep{Johnson1973}, while published observational data sets the bound of 18 km \citep{Carry2008}. More recently, images by the probe \textit{Dawn} also reveal an even smoother surface. Therefore, the association of Owings first contact to a relief is improbable.

@@

The 2013 occultation, therefore, offers not only an independent verification of the figures resulted from the 2010 event, but also validates the procedure carried out there which led to the best-constrained parameters in this work.

@@

\bibitem[\protect\citeauthoryear{Assafin et al.}{2011}] {2011gfun.conf...85A} Assafin M., et al., 2011, in Tanga P., Thuillot W., eds, Gaia follow-up network for the solar system objects: Gaia FUN-SSO workshop proceedings. IMCCE-Paris Observatory, Paris, p. 85

>>>>

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage[T1]{fontenc}
\usepackage{prettyref}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{color}
\usepackage[breaklinks=true]{hyperref}

@@

\title[Results of two stellar occultations by Ceres]{Results of two multi-chord stellar occultations by dwarf planet (1) Ceres}
\author[A. R. Gomes-J\'unior, B. L. Giacchini, F. Braga-Ribas et al.]{A. R. Gomes-J\'unior$^{1,}$\thanks{E-mail: altair08@astro.ufrj.br},
B. L. Giacchini$^{2,3,4}$,%\thanks{E-mail: breno@cbpf.br},
F. Braga-Ribas$^{5, 6, \dag}$
M. Assafin$^{1, \dag, \ddag}$,\newauthor
R. Vieira-Martins$^{1,5, \dag, \ddag}$,
J.I.B. Camargo$^{5, \dag}$,
B. Sicardy$^{7}$,
B. Timerson$^{4}$, 
T. George$^{4}$, \newauthor
J. Broughton$^{8}$, 
T. Blank$^{4}$,
G. Benedetti-Rossi$^{5}$, 
J. Brooks$^{4}$,  
R. F. Dantowitz$^{9}$,  \newauthor
D. W. Dunham$^{4}$, 
J. B. Dunham$^{4}$, 
C. K. Ellington$^{4}$,
M. Emilio$^{10}$,
F.R. Herpich$^{11}$,  \newauthor
C. Jacques$^{12}$,
P. D. Maley$^{4,13}$,
L. Mehret$^{10}$,
A.J.T. Mello$^{11}$,
A.C. Milone$^{14}$,\newauthor
E. Pimentel$^{12}$,  
W. Schoenell$^{11}$,
N. S. Weber$^{9}$
\\
$^{1}$Observat\'orio do Valongo/UFRJ, Ladeira Pedro Ant\^onio 43,
CEP 20.080-090 Rio de Janeiro - RJ, Brazil\\
$^{2}$Centro Brasileiro de Pesquisas F\'isicas, Rua Dr. Xavier Sigaud 150, Rio de Janeiro  22290-180, Brazil\\
$^{3}$Se\c{c}\~ao de Oculta\c{c}\~oes/REA-Brasil, Av. Prof. Magalh\~aes Penido 403/302, Belo Horizonte 31270-700, Brazil\\
$^{4}$International Occultation Timing Association, P.O. Box 131034, Houston, TX 77219-1034, USA\\
$^{5}$Observat\'orio Nacional/MCT, R. General Jos\'e Cristino 77, CEP 20921-400 Rio de Janeiro - RJ, Brazil\\
$^{6}$Federal University of Technology - Paran\'a (UTFPR / DAFIS), Rua Sete de Setembro, 3165, CEP 80230-901, Curitiba, PR, Brazil\\
$^{7}$LESIA, Observatoire de Paris, CNRS UMR 8109, Universit\'{e} Pierre et Marie Curie, Universit\'{e} Paris-Diderot, 5 place Jules Janssen,\\ F-92195 Meudon Cedex, France\\
$^{8}$RASNZ Occultation Section, P.O. Box 3181, Wellington, New Zealand\\
$^{9}$Clay Center Observatory at Dexter Southfield, 20 Newton Street, Brookline, MA 02445, USA\\
$^{10}$Universidade Estadual de Ponta Grossa, Ponta Grossa, Brazil\\
$^{11}$Universidade Federal de Santa Catarina, Florian\'opolis, Brazil\\
$^{12}$Centro de Estudos Astron\^omicos de Minas Gerais, Belo Horizonte, Brazil\\
$^{13}$NASA Johnson Space Center Astronomical Society, Houston, Texas, USA\\
$^{14}$Instituto Nacional de Pesquisas Espaciais, S\~ao Jos\'e dos Campos, Brazil\\
$^\dag$ Associated to Laborat\'{o}rio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, CEP 20921-400,\\ Rio de Janeiro, Brazil\\
$^\ddag$Affiliated researcher at Observatoire de Paris/IMCCE, 77 Avenue Denfert Rochereau 75014 Paris, France
}
\begin{document}

@@

\begin{abstract}
We report the results of two multi-chord stellar occultations by the dwarf planet (1) Ceres that were observed  from Brazil on 2010 August 17, and from the USA on 2013 October 25. Four positive detections were obtained for the 2010 occultation, and nine for the 2013 occultation. Elliptical models were adjusted to the observed chords to obtain Ceres' size and shape. Two limb fitting solutions were studied for each event. The first one is a nominal solution with an indeterminate polar aspect angle. The second one was constrained by the pole coordinates as given by \cite{Drummond2014}. Assuming a Maclaurin spheroid, we determine an equatorial diameter of $972 \pm 6~km$ and an apparent oblateness of $0.08 \pm 0.03$ as our best solution. These results are compared to all available size and shape determinations for Ceres made so far, and shall be confirmed by the NASA's \textit{Dawn} space mission.
\end{abstract}

@@

This work is organized as follows. In Sections 2 and 3 we analyse the 2010 and the 2013 occultations, respectively. In Section 4 we give the geocentric positions of Ceres derived from the occultations. The comparison of both results to those in the literature is carried out in Section 5.

@@

As predicted, on 2010 August 17 Ceres occulted the star TYC 6833-163-1 (UCAC4 313-111823), which has magnitude $V = 11.55$ and has the ICRS position, based on UCAC4 catalogue \citep{Zacharias2013}, to the date of occultation, corrected of proper motion:

%abaixo (UCAC2 20678706),
\begin{equation}
\left\{ 
  \begin{array}{l l}
    \alpha = 17^{\rmn{h}}18^{\rmn{m}}29\fs0085\\
    \delta = -27\degr 26\arcmin 38\farcs 867.
  \end{array}
\right.
\label{Eq: 2010-coord}
\end{equation}

@@

Observations were carried out in Brazil from five different sites as displayed in Table \ref{Tab: obs-2010} and Fig. \ref{Fig: Ceres-2010-map}. The occultation was detected from four of them. The southern-most one (UFSC) had a negative chord. From the positive ones, the one named INPE started observing after the event was already in progress, thus providing only the star's reappearance time; the other three recorded the whole phenomenon.

A remarkable circumstance of this event was the low velocity of Ceres: only $3.9~km~s^{-1}$ in the plane of the sky. Therefore, even exposures of a few seconds would translate in a relevant spatial resolution.

@@

\begin{table*}
 \centering
 \begin{minipage}{140mm}
  \caption{Circumstances of observation for all observing stations of the 2010 event. \label{Tab: obs-2010}}
  \begin{tabular}{@{}lccccc}
  \hline
          & Longitude & Telescope & Exposure & Result &   \\
     Site & Latitude  & Aperture  & Cycle time & Ingress & Observer\\          
          & Height    & f-ratio & Camera & Egress    & \\          
\hline
 Belo Horizonte & 43$\degr$59\arcmin 51.1\arcsec~W & LX200 & 5 s & Positive & C. Jacques  \\
 CEAMIG-REA &19$\degr$49\arcmin 49.0\arcsec~S & 31 cm & 12 s & 22:39:03.9 $\pm$ 0.6 s &  E. Pimentel \\
            & 825 m       &  f/10         &   SBIG ST10        & 22:40:20 $\pm$ 5 s &   \\
 & & & & & \\
 Pico dos Dias    & 45$\degr$34\arcmin45.1\arcsec~W &  Zeiss     & 1 s & Positive & J. I. B. Camargo \\
 LNA    &22$\degr$32\arcmin03.7\arcsec~S & 60 cm & 2 s & 22:37:30.3 $\pm$ 0.6 s &  G. B. Rossi \\
            & 1864 m     & f/12.5          &    Andor Ikon       & 22:41:55.3 $\pm$ 0.7 s &              \\
 & & & & & \\
 S\~ao Jos\'e dos       & 45$\degr$51\arcmin43.0\arcsec~W &  C11  & 2 s & Egress only & A. C. Milone\\
 Campos       &23$\degr$12\arcmin32.0\arcsec~S & 28 cm & 5 s & Start obs.: 22:39:44 & T. Maldonado\\
 INPE      & 620 m      & f/6.3          & SBIG ST7     & 22:42:03.0 $\pm$ 0.2 s & M. Okada    \\
 & & & & & \\
 Ponta Grossa       & 50$\degr$05\arcmin56.0\arcsec~W & RCX 400 &30 s & Positive & M. Emilio   \\
 UEPG       &25$\degr$05\arcmin22.2\arcsec~S & 40 cm & 52 s & 22:37:17 $\pm$ 13 s & L. Mehret   \\
            & 910 m      & f/8          & SBIG STL6E     & 22:39:56 $\pm$ 13 s & \\
 & & & & & \\
 Florian\'opolis       & 48$\degr$31\arcmin20.5\arcsec~W &   C11      & 3 s & No occultation  & W. Schoenell\\
 UFSC       &27$\degr$36\arcmin12.3\arcsec~S & 28 cm & 6 s&  Start obs.: 21:49:27 & A. J. T. Mello\\
            & 20 m       & f/10          &  SBIG ST7      &   End obs.: 22:51:21      & F. R. Herpich \\
\hline
\end{tabular}
\end{minipage}
\end{table*}

\subsection{Light curves and timing \label{Sec: light-curve-2010}}

@@

The ingress (disappearance) and egress (reappearance) instants of the occultation were obtained for each light curve by fitting a square-well model convoluted with the Fresnel diffraction, the CCD bandwidth, the stellar apparent diameter, and the applied finite exposure time; see \cite{Widemann2009} and BR13. %\cite{BragaRibas2013}.
%The Fresnel scale ($F = \sqrt{\lambda D/2}$) for the present Ceres' geocentric distance $D = 2.29 \text{ au} = 3.42 \times 10^{8}$~km is 0.3~km for a typical wavelength of $\lambda = 0.65$~$\micron$. The star diameter is estimated using the formulae of \cite{vanBelle1999}. Its $B$, $V$, and $K$ apparent magnitudes are 12.00, 11.78 and 9.18, respectively, in the Tycho catalog \citep{Hog2000}. This yields a diameter of about 0.11~km projected at Ceres' distance. 
The smallest integration time used in the positive observations was 1.0~s, which translates to almost 3.9~km in the celestial plane. Therefore, the time error of ingress and egress are largely dominated by the integration times, not by Fresnel diffraction or star diameter, which are on the order a few hundreds meters for this event.
The occultation data fit consists in minimizing a classical $\chi^{2}$ function for each light curve, as described in \cite{Sicardy2011} and BR13.\ The free parameter to adjust is the ingress or egress instant, which provides the minimum value of $\chi^{2}$, denoted as $\chi^{2}_{min}$. The best fittings to the 2010 occultation light curves are shown in Fig. \ref{Fig: Ceres-2010-curves}, and the derived occultation times are listed in Table~\ref{Tab: obs-2010}.
%The occultation fits consist in minimizing a classical $\chi^{2}$ function for each light curve, as described in \cite{Sicardy2011} and BR13.\ The free parameter to adjust is the ingress (disappearance) or egress (reappearance) time $t_{occ}$, which provides the minimum value of $\chi^{2}$, denoted as $\chi^{2}_{min}$. The best fittings to the 2010 occultation light curves are shown in Fig. \ref{Fig: Ceres-2010-curves}, and the derived occultation times are listed in Table~\ref{Tab: obs-2010}.

@@

We adopt an elliptic model for the limb profile, resulting from the projection of an oblate spheroid onto the sky plane. This choice is supported by the work of \cite{Drummond2014}, by means of direct imaging of Ceres. Hence, we have $N=7$ chord extremities to adjust the $M=5$ parameters which define an ellipse: apparent semimajor and semiminor axis ($a^\prime$ and $b^\prime$, respectively), position angle $P$ of its semiminor axis and the position $(f_c,g_c)$ of its centre with respect to the occulted star. 
Of course, the apparent semimajor axis {$a^\prime$} is equivalent to the equatorial radius $R_{equa}$ of the ellipsoid.
The coordinates $f_{c}$ and $g_{c}$, in kilometres, are calculated using the JPL\#33 Ceres' ephemeris \citep{Giorgini1996} and the occulted star's position. They are positive toward the local celestial east and north, respectively. The position angle $P$ is counted positively from the direction of local celestial north to celestial east. The apparent oblateness can be defined by $\epsilon^\prime = 1 - (b^\prime/a^\prime)$. The best-fitting solution is obtained minimizing a reduced $\chi^{2}$ function, where we define the number of degrees of freedom of the problem as $\mathcal{N} \equiv N - M$.  All the procedures that allow the determination of the error bars of the physical parameters can be found on BR13.

@@

Two possible solutions were considered for the limb fitting. The first is the nominal one, which consists of determining the five parameters that characterize an ellipse from the seven observed contacts. As we shortly show, it led to a rather large uncertainty on the position angle. Furthermore, the nominal solution alone is not capable of returning the true oblateness, which can be evaluated through equation 2 of BR13  provided that the aspect angle $\zeta$ is known.% or constrained by some physical hypothesis.

@@

\begin{table*}
 \centering
 \begin{minipage}{140mm}
  \caption{Results of limb fitting to the data of the 2010 and 2013 events.\label{Tab: results}}
  \begin{tabular}{@{}lcccc}
  \hline
     Solution & 2010/Nominal & \textbf{2010/Pole-constrained} & 2013/Nominal & 2013/Pole-constrained \\
\hline
Equatorial diameter (km) & 982 $\pm$ 14 & \textbf{972 $\pm$ 6}  & 971 $\pm$ 7  & 971 $\pm$ 7\\
%Equatorial radius (km) & 491 $\pm$ 7  & \textbf{486 $\pm$ 3}  & 485.5 $\pm$ 3.5 & 485.5 $\pm$ 3.5\\
True oblateness        & 0.08 $\pm$ 0.03 & \textbf{0.08 $\pm$ 0.03} & 0.08 $\pm$ 0.04 & 0.08 $\pm$ 0.04\\
Position angle (deg)   & 5 $\pm$ 10    & \textbf{12 $\pm$ 3} (*)& 22 $\pm$ 5    & 25 $\pm$ 3 (*)\\
$f_c$ (km)             & 133 $\pm$ 9   & \textbf{138 $\pm$ 5}   & 77 $\pm$ 6    & 78 $\pm$ 6\\
$g_c$ (km)             & -17 $\pm$ 15  & \textbf{-11 $\pm$ 11}  & 13 $\pm$ 16   & 13 $\pm$ 16\\
$\chi^2_{r,min}$       & 0.24          &  \textbf{0.42}         & 1.27          & 1.27\\
\hline
\end{tabular}
\textbf{Notes.} In bold we highlight our best solution. Error bars are at 1-sigma level. The polar diameter ($D_{pol}$) can be easily calculated from $D_{pol}=D_{equa}(1 - \epsilon)$.
(*) Position angle derived from Ceres' rotational pole coordinates determined by \cite{Drummond2014}.
\end{minipage}
\end{table*}

\begin{figure}
\includegraphics[scale=0.36]{figures/Ceres_2010_sphere.pdf} 
\caption{The best elliptical fit for the occultation chords for the event of 2010 using the times from Table~\ref{Tab: obs-2010} and the pole-constrained solution. The arrow indicates the direction of motion, blue lines are the observed chords, the red segments are the ingress, egress and mid-occultation error bars at 1-$\sigma$ level. \label{Fig:Ceres-2010-body}}
\end{figure}

@@

The event which took place on 2013 October 25 involved the star TYC 865-911-1 (UCAC4 496-058191), of magnitude $V = 10.05$. Based on UCAC4 \citep{Zacharias2013}, the ICRS position to the date of occultation corrected of proper motion is:
%
\begin{equation}
\left\{ 
  \begin{array}{l l}
    \alpha = 11^{\rmn{h}}57^{\rmn{m}}52\fs7641\\
    \delta = +09\degr 07\arcmin 49\farcs835
  \end{array}
\right.
\end{equation}
%
The occultation could only be visible in the United States, before dawn.% The shadow would then cross the Atlantic and Africa, but during daylight time.

@@

Nine positive chords were obtained by the variety of instruments listed in Table \ref{Tab: obs-2013}. Each station was equipped with a video camera with negligible readout times. This was of particular importance, since in this event Ceres' shadow speed was $42.6~ km~s^{-1}$, much faster than in the 2010 event. All the observing sites are in the United States, as depicted in Fig.~\ref{Fig: Ceres-2013-map}.

@@

\subsection{Light curves and timing}

@@

The immersion recorded at Owings also seems to be shifted (delayed) with respect to the nearby chords (see Fig. \ref{Fig: Ceres-2013-body}). This chord, actually, has roughly the same length of Mechanicsville's, despite the fact that they are separated by about $100~km$. Differently from Varina, though, this station had time stamps inserted in each frame of the video, which makes it more unlikely to justify an eventual timing issue. Another possible explanation for the apparent problem of the times of this chord is the determination of the ingress and egress instants in the light curve analysis, which could have been affected by noise. Finally, the delay during the ingress could be caused by a relief feature in Ceres; we shall soon return to this hypothesis.

@@

In the hypothesis that the delay observed in the immersion at Owings could be associated to a limb topography feature, the recorded contact would then correspond to a negative elevation of $31 \pm 4$ km with respect to the best-fitting ellipse. Theoretical models, however, predict that reliefs in Ceres should not be higher than about 10--20 km \citep{Johnson1973}, while published observational data sets the bound of 18 km \citep{Carry2008}. More recently, images by the probe \textit{Dawn} also reveal an even smoother surface. Therefore, the association of Owings first contact to a relief is improbable.

@@

The 2013 occultation, therefore, offers not only an independent verification of the figures resulted from the 2010 event, but also validates the procedure carried out there --and which led to the best-constrained parameters in this work.

@@

\bibitem[\protect\citeauthoryear{Assafin et al.}{2011}] {2011gfun.conf...85A} Assafin M. et al., 2011, Gaia follow-up network for the solar system objects : Gaia FUN-SSO workshop proceedings, held at IMCCE -Paris Observatory, France, November 29 - December 1, 2010. ISBN 2-910015-63-7
